### 2021-11-19 tables/table5.tex

% require package makecell 
\begin{table}[thb]\centering
    \caption{A table with multi-column headers and bold font highlights.}
    \label{tab:table5}
    \resizebox{0.48\textwidth}{!}{
    \large
    \begin{tabular}{c|c|ccc|ccc}
        \toprule
         %&  &  \multicolumn{3}{c}{\makecell{Static Scenes \\ w/ Ground Truth}} & \multicolumn{3}{c}{\makecell{Dynamic Scenes \\ w/o Ground Truth}} \\
        &  &  \multicolumn{3}{c|}{$6-9$ frames} & \multicolumn{3}{c}{$5-7$ frames} \\
        Data & Size &  \makecell{2-Exp \\ Break} &  \makecell{2-Exp \\ Break} &  \makecell{2-Exp \\ Break} &  \makecell{2-Exp \\ Break} &  \makecell{2-Exp \\ Break} &  \makecell{2-Exp \\ Break} \\
        \midrule
        A & $1280\times 720$ & 1 & 2 & 3 & 4 & \textbf{5} & \textbf{7} \\
        Ours & $4096\times 2168$ & \textbf{2} & \textbf{3} & \textbf{4} & \textbf{6} & \textbf{5} & 4 \\
        \bottomrule
    \end{tabular}
    }
\end{table}