### 2021-11-19 tables/table11.tex

\begin{table*}[t]
\begin{minipage}[c]{0.48\textwidth} \centering
    \caption{A table with caption at the right. This is useful for single-column paper (\eg, ECCV), where space are limited.} \label{tab:table11}
\end{minipage}\hfill
\begin{minipage}[c]{0.48\textwidth}\centering
    \resizebox{\textwidth}{!}{
    \large
    \begin{tabular}{*{10}{c}}
        \toprule
        &  &  \multicolumn{2}{c}{$6-9$ frames} & \multicolumn{2}{c}{$5-7$ frames} & \multicolumn{2}{c}{$50-200$ frames} \\
       Data & Size &  2-Exp & 3-Exp &  2-Exp & 3-Exp &  2-Exp & 3-Exp \\
        \midrule
        A & $1280\times 720$ & 1 & 2 & 3 & 4 & 5 & 4 \\
        Ours & $4096\times 2168$ & 2 & 3 & 4 & 6 & 5 & 4 \\
        \bottomrule
    \end{tabular}
    }
\end{minipage}
\end{table*}